---
title: "Loops, vectorized functions, and lists"
author: "Pablo Barbera"
date: "July 31, 2017"
output:
  html_document: default
  pdf_document: default
---

Being designed for statistics and data analysis, `R` has powerful built-in functions for data manipulation. However, you can dramatically extend `R`'s functionality by writing your own functions.

### Making functions

`R` functions are objects just like the vectors and data frames we've worked with, so we create them using an assignment.

```{r}
times_2 <- function(x) x * 2
times_2(6)
times_2(1:5)
```

For longer functions, it's necessary to use curly braces `{}`. We can also input multiple objects into a function, and return more complex objects, such as a vector or list.

```{r}
two_numbers <- function(x, y) {
  my_sum <- x + y
  my_product <- x * y
  my_ratio <- x / y
  return(c(my_sum, my_product, my_ratio))
}
two_numbers(4, 11.93)
```

### Loops

"For" loops are probably the most common type of loop and are easily implemented in `R`. For instance, we can get the first 40 terms in the Fibonacci sequence using a for loop.

```{r}
fib <- c(0, 1, rep(NA, 38)) # initialize fib sequence
for(i in 3:40) {
  fib[i] <- fib[i-1] + fib[i-2]
}
```

### If statements

Depending on whether a condition is true or false, we might want to execute different chunks of code.

```{r}
compare_xy <- function(x, y) {
  if (x < y) {
    print("y is greater than x")
  } else if (x > y) {
    print("x is greater than y")
  } else {
    print("x and y are equal")
  }
}
compare_xy(3, 4)
compare_xy(4, 3)
compare_xy(1, 1)
```

