---
title: "Challenge 8: Analyzing the Amnesty International Annual Report"
author: Pablo Barbera
date: "August 3, 2018"
output: html_document
---

In this challenge you'll put your PDF-parsing skills to practice by analyzing the most recent [annual report](https://www.amnesty.org/en/latest/research/2017/02/amnesty-international-annual-report-201617/) from Amnesty International on the state of Human Rights in 159 countries.

Pick a country and use `pdftotext` to convert the text from the specific pages of the report that refer to that country.

```{r}

```

Read the .txt file into R and use a combination of regular expressions and string parsing functions to remove the text that is not part of the report that corresponds to that country.

```{r}

```

Collapse the text into a single character vector and, using the text analysis techniques we learned earlier in the workshop, produce a word cloud of the most common words in the report.

```{r}

```

(Optional) Parse the entire text of the report, and then extract the text referring to each country and create a data frame with the length (in words) of each country report. [Clue: try to find something that helps you "anchor" the beginning of each country.] Which is the country with the longest report?


```{r}


```



