---
title: "Challenge 7"
author: Pablo Barbera
date: August 2nd, 2018
output: html_document
---

Turning to the REST API, collect the 2,000 most recent tweets from a Twitter account that you find relevant.

```{r}

```

Using the code we just used, look at the most popular hashtags used by that account. What do you learn?

```{r}


```

Produce a word cloud of this set of tweets to look at what are the words most frequently used by this user.

```{r}

```

Estimate the political ideology of this user.

```{r}

```

Then, find the five screen names most frequently mentioned by this user and estimate their political ideology too. Are they all similar? (Note that this method will only work for US-based accounts that do not follow a lot of other users.)

```{r}

```

Download the followers of that account (if the account has more than 5K-10K followers, choose another one that has a number below that limit.) Take a small random sample of followers (maybe 10-20) and run the botometer. What proportion of followers do you estimate to be bots?


